
```{r load libraries, set projection}
rm(list=ls())
library(sp)
library(rgdal)
library(rgeos)
library(maptools)  
library(raster)

basedir<-"C:/ALR"
proj4.NHD<-"+proj=longlat +ellps=GRS80 +datum=NAD83 +no_defs"

```


```{r load spatial high rez vector data}
## load high rez vector polygon data
setwd(file.path(basedir,"Data/StreamData/custom_delineated_CT"))
watersheds<-readShapePoly("Watershed_prj",proj4string=CRS(proj4.NHD))
watersheds.clipped<-readShapePoly("Watershed_prj_clip",proj4string=CRS(proj4.NHD))
catchments<-readShapePoly("Catchment_prj",proj4string=CRS(proj4.NHD))

names(watersheds)
names(catchments)
nrow(watersheds@data)
nrow(catchments@data)

```

```{r check for unique watersheds/catchments}


```

```{r duplicate spatial objects?}
# watersheds2<-watersheds[,"HydroID"]

# row.names(watersheds2@data)<-watersheds2$HydroID
# catchments2<-catchments[,"HydroID"]
# row.names(catchments2@data)<-catchments2$HydroID

# row.names(watersheds@data)<-NULL
# head(watersheds)
# row.names(catchments@data)<-NULL
# head(catchments)
# class(catchments)
```

```{r find the catchments w/in each watershed and save to a file}

setwd(file.path(basedir,"/Models/zonal_stats_high_rez"))

# for (i in 1:nrow(watersheds2))
# j<-21217
# testbatch<-watersheds@data[20:200,"HydroID"]
# first<-TRUE
# for (j in testbatch) {

for (j in watersheds$HydroID) {
  c<-c()
  i<-which(watersheds$HydroID==j)
  print(i)   
  w<-watersheds[i,]
  print(w@data)
  # c1<-w %over% catchments2
  wbuff<-gBuffer(w,width=-.001,id="HydroID")
  {
#     if (catchments$HydroID==w$HydroID) {
      #if the catchment is the watershed
    if (all.equal.numeric( as.vector(coordinates(w)),
                           as.vector(coordinates(catchments[catchments$HydroID==w$HydroID,])),
                           tolerance=0.0001)==T) {
      #if the original (not buffered) waterhsed coordinates are identical to the 
      #   downstream catchment (the one w/ a matching HydroID)
      #   (the catchment is the watershed)
#       print(paste("w:",w$HydroID,"c:",catchments@data[i,"HydroID"]))
      c<-w$HydroID
      print("catchment is watershed")
      print( paste("w:",coordinates(w),"   c:",coordinates(catchments[catchments$HydroID==w$HydroID,])))
    }
    else {
      #get list of all catchments w/in buffered (contracted) watershed
      tryCatch(
        c1<-over(wbuff,catchments,returnList=T),
        #If there are no catchments completely contained by the contracted watershed 
        #   because they're outside the extent of the catchments shapefile,
        #   "over" function will return error
        #Instead, say that the watershed consists of only the downstream catchment
        #   (the catchment w/ the same HydroID as the watershed)
#         with the s 
        error=function(e) {
          c<-j
        }
      )
      #save their HydroIDs
      if (length(c)>0) #if error wasn't generatred above
        c<-catchments@data[c1[[1]],"HydroID"]
    }
  }
  print(paste("w:",w$HydroID))
  print(c)


  
#   to.write<-c(watersheds2@data[i,],c)
#   print(to.write)
  write.table(matrix(c(w$HydroID,c),nrow=1),"watershed_catchmentIDs.csv",append=T,quote=F,sep=",",row.names=F,col.names=F)
#   first<-FALSE

  }
```

```{r}
  #check that the farmington values are the same from catchment and watershed
  
  #next, look at a *different* shapefile, of watersheds clipped to exclude catchment outline
  #if the hydroid is in there, it means that part of the watershed is not included in the catchments we just selected
  #if that is true, do a buffer of the cropped watershed, overlap that buffer with a shapefile of *cropped* huc12s
  #write the huc12 ids in a separete csv, with the watershed hydroid in the first column
  
  #when reassembling the watershed zonal stats with spatial average, need to pull the attributes of two separate files:
  #one of the catchments
  #and one of the cropped hucs
  
  #gis work to do:
  #create outline of catchments
  #clip watersheds to exclude catchments outline
  #clip huc12s to exclude catchments outline
  #run zonal stats on non-overlapping catchments
  #run zonal stats on non-overlapping clipped HUC12s
  #     )
```


```{r}
watersheds2@data[i,]
c2<-over(w,catchments2,returnList=T)
c2

plot(catchments2[c2[[1]],],border="blue",add=F)
plot(w,col="yellow",add=T)
plot(catchments2[c2[[1]],],border="blue",add=T)
plot(wbuff,add=T)

test.ids<-read.csv("C:/Users/arosner/Desktop/farm_wshed_check.txt")
head(test.ids)

check<-0
for (i in 1:2) {
  #   print(  if (!all.equal( coordinates(w)[i],coordinates(catchments2[catchments2$HydroID==w$HydroID,])[i],tolerance=0.001)))
  if (!all.equal( coordinates(w)[i],coordinates(catchments2[catchments2$HydroID==w$HydroID,])[i],tolerance=0.001))
    check<-1
}
check



all.equal.numeric( as.vector(coordinates(w)),as.vector(coordinates(catchments2[catchments2$HydroID==21208,])),tolerance=0.0001)==T




print(all.equal( coordinates(w),coordinates(catchments2[catchments2$HydroID==w$HydroID,]),tolerance=0.001))

```

```{r load spatial raster data}

```






setwd("C:/ALR/occupancy_model_CT/zonal_stats_overlap")

setwd
i<-watersheds2$HydroID==21886
which(i)


class(coordinates(w)[i])
class(coordinates(catchments2[catchments2$HydroID==w$HydroID,])[i])

# plot(catchments2,add=T,border="blue")


# plot(catchments2[which(catchments2$HydroID == c1[[1]]),])
plot(catchments2[c1[[1]],],col="yellow")
# plot(catchments2[c2,],border="red",add=T)
plot(wbuff,border="blue",add=T)
# plot(wbuff,add=T,border="red")






# i<-which(watersheds2$HydroID==23186)
# i
w<-watersheds2[watersheds2$HydroID==23186,]
# c1<-w %over% catchments2
wbuff<-gBuffer(w,width=-.001,id="HydroID")
c1<-over(w,catchments2,returnList=T)
c1<-as.vector(c1$HydroID)
c2<-over(wbuff,catchments2,returnList=T)
c2<-as.vector(c2$HydroID)
plot(catchments2[c1,],col="yellow")
plot(catchments2[c2,],border="red",add=T)
plot(wbuff,border="blue",add=T)
# plot(wbuff,add=T,border="red")
c1
c2
wbuff
# c2<-gContains(w,catchments2)
# 1519 %in% c3
# c<-catchments[,]
dim(c2)
class(c2)
# plot(catchments[catchments$,])
plot(watersheds2[33,],col="red")
plot(catchments2,border="blue",add=T,col=NA)
plot(catchments2[c,],border="yellow",add=T)

# over2<-function(outline,subsections) {
#   buf<-gBuffer(outline,width=-1)
#   over(buf,subsections,returnList=T)
# }


# cc<-gCentroid(catchments2[c3[1],])
# for (i in 1:length(c3)) 
#   cc[i,]<-gCentroid(catchments2[c3[i],])
# 
# plot(cc,col="blue",add=T)
plot(watersheds2[33,],border="red",add=F)
plot(catchments2[c2,],border="blue",add=T)

plot(catchments2[c2,],border="blue",add=F)
plot(catchments2[c3,],border="blue",add=F)

over()





###
### load and reclassify raster data
###
raster.index<-c()
raster.border<-c()

impervious<-raster("C:/ALR/Data/LandData/NLCD_land_cover/clipped/imp_ctprj_nul",crs=CRS(proj4.NHD))
road.crossings<-raster("C:/ALR/Data/LandData/TIGER_roads/cross_density",crs=CRS(proj4.NHD))
road.density<-raster("C:/ALR/Data/LandData/TIGER_roads/road_density",crs=CRS(proj4.NHD))
raster.index<-c(raster.index,"impervious","road.crossings","road.density")
raster.border<-c(raster.border,"state","state","state")



precip<-raster("C:/ALR/Data/ClimateData/PRISM/CT_raster/annual2/prcpann_r_prj",crs=CRS(proj4.NHD))
tmax<-raster("C:/ALR/Data/ClimateData/PRISM/CT_raster/annual2/tmax_r_prj",crs=CRS(proj4.NHD))
tmin<-raster("C:/ALR/Data/ClimateData/PRISM/CT_raster/annual2/tmin_r_prj",crs=CRS(proj4.NHD))
precip.nov<-raster("C:/ALR/Data/ClimateData/PRISM/CT_raster/annual2/prcp_1nov_prj",crs=CRS(proj4.NHD))
precip.winter<-raster("C:/ALR/Data/ClimateData/PRISM/CT_raster/annual2/prcp_win_prj",crs=CRS(proj4.NHD))
raster.index<-c(raster.index,"precip","tmax","tmin","precip.nov","precip.winter")
raster.border<-c(raster.border,"state","state","state","state","state")


fws.wetland<-raster("C:/ALR/Data/LandData/US FWS/processed/wetland",crs=CRS(proj4.NHD))
fws.water<-raster("C:/ALR/Data/LandData/US FWS/processed/open_water",crs=CRS(proj4.NHD))
raster.index<-c(raster.index,"fws.wetland","fws.water")
raster.border<-c(raster.border,"watershed","watershed")

glac.sed.thickness<-raster("C:/ALR/Data/LandData/CT DEEP/processed/glacsed_thick",crs=CRS(proj4.NHD))
surf.coarse<-raster("C:/ALR/Data/LandData/CT DEEP/processed/surf_coarse",crs=CRS(proj4.NHD))
drain.class<-raster("C:/ALR/Data/LandData/USDA_NRCS/processed/drn_class",crs=CRS(proj4.NHD))
sand<-raster("C:/ALR/Data/LandData/USDA_NRCS/processed/prc_sand",crs=CRS(proj4.NHD))
raster.index<-c(raster.index,"glac.sed.thickness","surf.coarse","drain.class","sand")
raster.border<-c(raster.border,"state","state","state","state")
# raster.index<-c(raster.index,"glac.sed.thickness","surf.coarse","drain.class","sand")
# raster.border<-c(raster.border,"state","state","state","state")


no3<-raster("C:/ALR/Data/DepositionData/CT/no3_ctwsd_prj",crs=CRS(proj4.NHD))
so4<-raster("C:/ALR/Data/DepositionData/CT/so4_ctwsd_prj",crs=CRS(proj4.NHD))
raster.index<-c(raster.index,"no3","so4")
raster.border<-c(raster.border,"watershed","watershed")

slope<-raster("C:/ALR/Data/TopoData/slope_prj",crs=CRS(proj4.NHD))
elevation<-raster("C:/ALR/Data/TopoData/dem_prj",crs=CRS(proj4.NHD))
raster.index<-c(raster.index,"slope","elevation")
raster.border<-c(raster.border,"watershed","watershed")

lcover.raster<-raster("C:/ALR/Data/LandData/NLCD_land_cover/clipped/lu_ctwshd_prj",crs=CRS(proj4.NHD))
lcover.table<-read.csv("C:/ALR/Data/LandData/LU_classifications/nlcd_classes_and_categories.csv")


#reclassify land cover in various ways, to create binary rasters of different categories
names(lcover.table)
# categories<-names(lcover.table)[3:length(names(lcover.table))]
categories<-names(lcover.table)[3:8]  #decided not to include wetland and water from nlcd, will use FWS wetland inventory instead
categories2<-gsub("_",".",categories)
categories2<-paste0("nlcd.",categories2)
categories
categories2

setwd("C:/ALR/Data/LandData/NLCD_land_cover/clipped/reclassified")
start.time<-proc.time()[3]
for (i in 1:length(categories)) {
  assign(categories2[i],subs(lcover.raster,lcover.table,which=categories[i],subsWithNA=F))
  print(paste0(categories2[i],": ",(proc.time()[3]-start.time)/60))
  writeRaster(get(categories2[i]),paste0("nlcd_",categories[i],".tif"),"GTiff")
}
(proc.time()[3]-start.time)/60

xforest<-raster("forest_prj",crs=CRS(proj4.NHD))
plot(xforest)

raster.index<-c(raster.index,categories2)
raster.border<-c(raster.border,rep("watershed",length(categories2)))

# writeRaster(get(categories2[i]),paste0(categories[i]),"raster")


raster.index
raster.border


# by.HydroID<-watersheds@data[order(watersheds@data$HydroID),]
# by.HydroID$index<-row.names(by.HydroID)
# #just checking to make sure that the two shapefiles are in the same order, as they should be
# by.HydroID2<-watersheds.clipped@data[order(watersheds.clipped@data$HydroID),]
# by.HydroID2$index<-row.names(by.HydroID2)
# test<-by.HydroID$HydroID-by.HydroID2$HydroID
# sum(test)









### 
### start zonal statistics 
###

setwd("C:/ALR/occupancy_model_CT/zonal_stats_overlap")

### this creates an empty data.frame
zonal.stats<-data.frame(matrix(ncol=length(raster.index)+3,nrow=0)) 
names(zonal.stats)<-c("HydroID",raster.index,"elev.max","elev.min")
zonal.stats
errors1<-c()
errors2<-c()
errors3<-c()
# errors.raster1<-data.frame(HydroID=integer(),ii=integer(),stringsAsFactors=F)
# errors.raster2<-data.frame(HydroID=integer(),ii=integer(),raster.layer=character(),stringsAsFactors=F)
# errors.zonal<-data.frame(HydroID=integer(),ii=integer(),stringsAsFactors=F)

# ii<-0

### big loop (though watersheds) ###
start.time<-proc.time()[3]
# for (i in 1:nrow(watersheds)) {
# for (i in by.HydroID$index) {
#   i<-as.character(i)
#   ii<-ii+1
for (i in 1:nrow(watersheds)) {
  #   i<-as.character(by.HydroID$index[ii])
  assign("flag1",0,envir=.GlobalEnv)
  
  
  
  #
  ### create raster of current watershed from poylgon
  #
  r<-raster(ncol=180, nrow=180) #create a generic raster
  tryCatch(
    extent(r)<-extent(watersheds[i,]), #set generic raster's extent to that of current watershed feature
    error=function(err) {
      print(paste0("error in resample, skip watershed ",watersheds@data[i,"HydroID"]," (",i,") column ",raster.index[j]," extent ",raster.border[j],", --- ",err))
      assign("flag1",1,envir=.GlobalEnv)
    })
  tryCatch(
    wshed.raster.watershed<-rasterize(x=watersheds[i,],r,field="HydroID",silent=T), #rasterize current feature, 
    error=function(err) {
      print(paste0("error in resample, skip watershed ",watersheds@data[i,"HydroID"]," (",i,") column ",raster.index[j]," extent ",raster.border[j],", --- ",err))
      assign("flag1",1,envir=.GlobalEnv)
    })
  tryCatch(
    extent(r)<-extent(watersheds.clipped[i,]), #set generic raster's extent to that of current watershed feature
    error=function(err) {
      print(paste0("error in resample, skip watershed ",watersheds@data[i,"HydroID"]," (",i,") column ",raster.index[j]," extent ",raster.border[j],", --- ",err))
      assign("flag1",1,envir=.GlobalEnv)
    })
  tryCatch(
    wshed.raster.state<-rasterize(x=watersheds.clipped[i,],r,field="HydroID",silent=T), #rasterize current feature
    error=function(err) {
      print(paste0("error in resample, skip watershed ",watersheds@data[i,"HydroID"]," (",i,") column ",raster.index[j]," extent ",raster.border[j],", --- ",err))
      assign("flag1",1,envir=.GlobalEnv)
    })
  #if failed to create raster from polygon
  if (flag1 ==1 ) {
    assign(errors1,c(errors1,watersheds@data[i,"HydroID"]))
    #     assign(errors.raster1[nrow(errors.raster)+1,]<-c(watersheds@data[i,"HydroID"]),ii)
  }
  
  #
  ### crop
  #
  if (flag1 ==0 ) { #(didn't fail when creating raster from polygon), otherwise, skip everythign for this watershed
    for (j in 1:length(raster.index)) {
      assign("flag2",0,envir=.GlobalEnv)
      assign("flag3",0,envir=.GlobalEnv)
      
      #for raster layers that only have data w/in the state border, use the watershed features clipped to buffered border of CT
{if (raster.border[j]=="state")
  wshed.raster<-wshed.raster.state
 else
   wshed.raster<-wshed.raster.watershed
      }
      
      #crop the base raster to the extents of the current feature
      #snap=out means that it will include cells that intersect the border
      #which also means it won't be the exact same size as the extent specified
      #       print(paste("before crop,",raster.index[j]))
      tryCatch(
        base<-crop( get(raster.index[j]), extent(wshed.raster),snap="out"),
        error=function(err) {
          print(paste0("caught error in crop ",watersheds@data[i,"HydroID"]," (",i,") column ",raster.index[j]," border ",raster.border[j],", --- ",err))
          base0=extend( get(raster.index[j]), extent(watersheds[i,]))
          tryCatch(
            assign("base",crop( base0, extent(wshed.raster),snap="out"),envir=.GlobalEnv),
            error=function(err) {
              print(paste0("error in crop not resolved by extend command, skip watershed ",watersheds@data[i,"HydroID"]," (",i,") column ",raster.index[j]," border ",raster.border[j],", --- ",err))
              assign("flag2",1,envir=.GlobalEnv)                
            })
        }
      )    
      
      #failed for this particular raster layer (and this particular watershed) to crop
      if (flag2 ==1 ) {
        #           assign(errors2[nrow(errors2)+1,],c(watersheds@data[i,"HydroID"],ii,raster.index[j]))
        assign("errors2",c(errors2,watersheds@data[i,"HydroID"]))
      }
      
      
      #
      ### resample
      #
      
      if (flag2 == 0) { #didn't fail when cropping/resampling this particular raster/watershed
        #match extents
        #this is necessary because since we used snap="out", base is now slightly larger than wshed extent to include edges
        #and it rounds off w/ slightly different decimal points
        tryCatch(
          extent(wshed.raster)<-extent(base), 
          error=function(err) {
            print(paste0("error in extent, skip watershed ",watersheds@data[i,"HydroID"]," (",i,") column ",raster.index[j]," extent ",raster.border[j],", --- ",err))
            assign("flag3",1,envir=.GlobalEnv)
          }) 
        
        #match resolutions and crop again
        #resolutions must match for zonal function
        #this changes the base resolution to match the selected feature resolution, 
        #and crops it exactly to the feature's extent
        tryCatch(
          base2<-resample(base,wshed.raster,method="ngb"), 
          error=function(err) {
            print(paste0("error in resample, skip watershed ",watersheds@data[i,"HydroID"],", column ",raster.index[j]," extent ",raster.border[j],", --- ",err))
            assign("flag3",1,envir=.GlobalEnv)
          }) 
        #failed for this particular raster layer (and this particular watershed) to resample
        if (flag3 ==1 ) {                           
          assign("errors3",c(errors3,watersheds@data[i,"HydroID"]))
        }          
        
        #calculate zonal statistics
        
        
        if (flag3==0 ) { #didn't fail for this particular raster/watershed to crop or resample
{
  if (j == 1 )
    zonal.stats[i,c("HydroID",raster.index[j])]<-zonal(base2, wshed.raster, stat='mean', na.rm=TRUE) 
  else 
    zonal.stats[i,raster.index[j]]<-zonal(base2, wshed.raster, stat='mean', na.rm=TRUE)[2] 
}
#get hydroid and mean of base raster layer 
if (raster.index[j]=="elevation") {
  zonal.stats[i,"elev.max"]<-zonal(base2, wshed.raster, stat='max', na.rm=TRUE)[2] 
  zonal.stats[i,"elev.min"]<-zonal(base2, wshed.raster, stat='min', na.rm=TRUE)[2]             
}
        }#end third flag check
        
      }#end second flag check
      
      
      rm(list=c("base","base2","base0"))
    }#end j loop, through raster columns  
  }#end first check for flag  
  
  
  
  #   print(paste0("Completed, not saved through i ",i, ", ii ", ii, ", HydroID ", watersheds@data[i,"HydroID"]))
  #to save processing time, this writes to a file every 25th feature
  #could write to file every time, or only at the end
  if ( i>25 & (i %% 25) == 1) {
    is.first<-((i-25)==1)
    write.table(zonal.stats[c((i-25):(i-1)),],"zonal_stats_5_10.csv",append=(!is.first),quote=F,sep=",",row.names=F,col.names=is.first)
    write.table(errors1,"errors1.csv")
    write.table(errors2,"errors2.csv")
    write.table(errors3,"errors3.csv")  
    print(paste0(round((proc.time()[3]-start.time)/60,digits=2), " minutes. Feature #",i,", HydroID ",watersheds@data[i,"HydroID"]))
  }
  
  rm(list=c("wshed.raster.watershed","wshed.raster.state","wshed.raster","r"))
  
}#end i loop, through watersheds

end.time<-proc.time()[3]
print(paste0("Time elapsed: ",round((end.time-start.time)/60,digits=2), " minutes"))

print(errors1)
print(errors2)
print(errors3)

# list(compData, as.list(c(5,443)x<-data.frame(a=0,b="jhkh")))
# library(data.table)
# x<-data.frame(a=integer(),b=integer(),c=character(),stringsAsFactors=F)
# x[nrow(x)+1,]<-c(5234,34,"sdfd")
# x
# 
# x<-rbindlist(list(x, list(a=5,b="sdfs")))
# x
